---
title: "Brobdingnagian matrices"
author: "Robin K. S. Hankin"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{brobmat}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r set-options, echo = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "#>", dev = "png", fig.width = 7, fig.height = 3.5, message = FALSE, warning = FALSE)
options(width = 80, tibble.width = Inf)
```

# Brobdgingnagian matrices

R package `Brobdingnag` has basic functionality for matrices.  It
includes matrix multiplication and addition, but determinants and
matrix inverses are not implemented.  First load the package:

```{r}
library("Brobdingnag")
```

The standard way to create a Brobdgingnagian matrix (a `brobmat`) is
to use function `brobmat()` which takes arguments similar to
`matrix()` and returns a matrix of entries created with `brob()`:

```{r}
M1 <- brobmat(-10:13,4,6)
colnames(M1) <- state.abb[1:6]
M1
```

Function `brobmat()` takes an argument `positive` which specifies the sign:

```{r}
M2 <- brobmat(
c(1,104,-66,45,1e40,-2e40,1e-200,232.2),2,4,
positive=c(T,F,T,T,T,F,T,T))
M2
```

Standard matrix arithmetic is implemented, thus:

```{r}
rownames(M2) <- c("a","b")
colnames(M2) <- month.abb[1:4]
M2
M2[2,3] <- 0
M2
M2+1000
```

We can also do matrix multiplication, although it is slow:


```{r}
M2 %*% M1
```


## Numerical verification: matrix multiplication

We will verify matrix multiplication by carrying out the same
operation in two different ways.  First, create two largish
Brobdingnagian matrices:

```{r}
nrows <- 11
ncols <- 18
M3 <- brobmat(rnorm(nrows*ncols),nrows,ncols,positive=sample(c(T,F),nrows*ncols,replace=T))
M4 <- brobmat(rnorm(nrows*ncols),ncols,nrows,positive=sample(c(T,F),nrows*ncols,replace=T))
M3[1:3,1:3]
```

Now calculate the matrix product by coercing to numeric matrices and
multiplying:q


```{r}
p1 <- as.matrix(M3) %*% as.matrix(M4)
```

and then by using Brobdingnagian matrix multiplying, and then coercing
to numeric:

```{r}
p2 <- as.matrix(M3 %*% M4)
```

The difference:

```{r}
max(abs(p1-p2))
```

is small.  Now the other way:

```{r}
q1 <- M3 %*% M4
q2 <- as.brobmat(as.matrix(M3) %*% as.matrix(M4))
max(abs(as.brob(q1-q2)))
```


## Numerical verification: integration with the `cubature` package

The matrix functionality of the `Brobdingnag` package was originally
written to leverage the functionality of the `cubature` package.  Here
I give some numerical verification for this.

Suppose we wish to evaluate

\[
\int_{x=0}^{x=4}(x^2-4)\,dx
\]

using numerical methods.  See how the integrand includes positive and
negative values; the theoretical value is $\frac{16}{3}=5.33\ldots$.
The `cubature` idiom for this would be

```{r,label = numericalintegration}
library("cubature")

f.numeric <- function(x){x^2 - 4}

out.num <- cubature::hcubature(f = f.numeric, lowerLimit = 0, upperLimit = 4, vectorInterface = TRUE)
out.num
```

and the Brobdingnagian equivalent would be


```{r,label = numericalintegrationbrob}

f.brob <- function(x) {
	x <- as.brob(x[1, ])
	as.matrix( brobmat(x^2 - 4, ncol = length(x)))
}

out.brob <- cubature::hcubature(f = f.brob, lowerLimit = 0, upperLimit = 4, vectorInterface = TRUE)
out.brob
```

We may compare the two methods:

```{r,label=comparebrobandnumeric}
out.brob$integral - out.num$integral
```
